---
title: "EDA-LA01 / Student_Test_Scores_Prediction"
author: "Bharat(1NT18IS039) & Adarsh(1NT18IS011)"
date: "11/07/2021"
output: html_document
---
# github link:"https://github.com/1NT18IS011-AdarshHegde/EDA-CE"


# Reading csv dataset file from local storage and storing it as dataset 
```{r}
dataset <- read.csv("D:/Academic/Sem VI/EDA-LA01/EDA-LA01/test_scores.csv")
```
# Printing the dataset
```{r}
dataset
```
# Getting structure of dataset
```{r}
str(dataset)
```
# Getting class of dataset
```{r}
class(dataset)
```
# Getting all the attributes of dataset
```{r}
ls(dataset)
```
# Printing values of dataset attributes
```{r}
head(dataset$gender,n=10)
```
```{r}
head(dataset$lunch,n=30)
```
# Getting summury of the dataset and attributes of dataset
```{r}
summary(dataset)
```
```{r}
summary(dataset$gender)
```
```{r}
summary(dataset$pretest)
```
# Length of the dataset and its attributes
```{r}
length(dataset)
```
```{r}
length(dataset$posttest)
```
```{r}
length(dataset$lunch)
```
# Getting column names of the dataset
```{r}
colnames(dataset)
```
# Coverting datatype of the Dataset attributes
```{r}
# converting character to integer
numrc=as.numeric(dataset$school)
head(numrc)
```
```{r}
# converting integer to character
numrch=as.character(dataset$pretest)
head(numrch)
```
# Saving history
```{r}
history=savehistory(file ='EDA-LA1.Rhistory')
```
# Show history
```{r}
history(show.max=10)
```
# Saving data files to disk
```{r}
save(dataset,file='C:/Users/Aadarsh Hegde/Downloads/La1.RData')
load('C:/Users/Aadarsh Hegde/Downloads/La1.RData')
```
# Manipulating vectors of dataset
```{r}
data1=dataset$posttest     #Assigning reading score values to data1
```
```{r}
head(data1,n=20)                           #reading values in the data1
```
```{r}
length(data1)                   #Getting length of data1
```
```{r}
max(data1)                      #Maximum value in the data1
```
```{r}
which(data1==max(data1))        #Getting positions of the occurrence of the maximum value of data1 
```
# Returning logical values from vector
```{r}
which(data1==77)
```
# Reading values with skipping values by particular number
```{r}
# Syntax : sequence(starting point , length , no.of interval)
dataset$n_student[seq(1,length(dataset$n_student),10)]
```
```{r}
dataset$pretest[seq(1,length(dataset$pretest),20)]
```
# Performing 'order' on dataset
```{r}
head(order(dataset),n=100)
```
```{r}
head(order(dataset$posttest),n=100)
```
#Performing 'sort' on dataset
```{r}
head(sort(dataset$gender),n=30)              #sorting gender of dataset
```
```{r}
head(sort(dataset$n_student),n=50)          #sorting of students
```
```{r}
sort(dataset[1,])                 #sorting on 1st row and all the columns of dataset
```
```{r}
sort(dataset[,1])                 #sorting on 1st column and all rows of dataset
```
# Performing 'rank' on dataset
```{r}
head(rank(dataset$student_id),n=50) #rank on test course preparation
```
```{r}
head(rank(dataset$school_type),n=50)              #rank on math score

```
# Manipulating dataframes
```{r}
dataset[3,3]                    #Reading values from 3rd row,3rd column
```
```{r}
dataset[4,]                     #Reading values from 4th row,all columns
```
```{r}
dataset[,2]                     #Reading values from all rows with only 2nd column
```
```{r}
dataset[3,1:4]                  #select the third row and display columns one to four
```
```{r}
dataset[c(1,5,10,15),]          #Specify several rows but leave out a value at the end to display all columns
```
```{r}
dataset[c(1,3,5,7),-4]               #specify several rows but display all columns except the fourth
```

# Quick looks at Complicated Data Objects
```{r}
# Reading first six values of dataset
head(dataset)
```
```{r}
# Reading last six values fo dataset
tail(dataset)
```
```{r}
# Reading desired number of values from the top or bottom
head(dataset,n=20)
```
```{r}
head(dataset$school,n=20)
```
```{r}
tail(dataset,n=10)
```
```{r}
tail(dataset$lunch,n=10)

```
# Viewing and Setting Names
```{r}
names(dataset)                  #Getting all column names of dataset
```
```{r}
head(row.names(dataset),n=20)              #Getting rows of dataset
```
```{r}
head(dimnames(dataset),n=50)               #Getting both rows and columns

```
# Rotating data tables
```{r}
dataset2=dataset
dataset2.t=t(dataset2)
head(dataset2.t,n=60)
```
```{r}
data2=head(dataset$posttest,n=10)
post.score.t=t(data2)
post.score.t

```
# Re-ordering data Frames
# Create an index using the values in the gender column, with ties resolved by the math.score column
```{r}
data2=with(dataset, order(gender,n_student))
head(data2,n=50)
```
# Converting dataframe into matrix
```{r}
data_matrix=as.matrix(dataset)
head(data_matrix,n=50)
```
# To add new column to data frame
```{r}
library(dplyr)
dataset2%>%mutate(newcol = NA)
```
# Getting a Subset of a Data Frame
```{r}
# Loading the package
library(gcookbook)                  
```
# To keep rows where gender is "male" and where the maths score is inclusive of and between 60 and 100. using filter()  function from gcookbook.
```{r}
dataset2 %>% filter(gender=="Male",posttest==60)
```
# Changing the Order of Factor Levels
```{r}
# Installing forcats package
library(forcats)
```
# To use two vectors, one with the original levels and one with the new ones, use do.call() with fct_recode().
```{r}
data3mappings <- setNames(dataset$school_setting,dataset$n_student)
head(data3mappings)

```
# Removing Unused Levels from a Factor
```{r}
head(fct_drop(dataset2$gender),n=20)
```
# Work on a subset of the dataset data set
```{r}
pg <- dataset[c(1:10),]
pg
```
# Calculating New Columns from Existing Columns
# To calculate a new column of values in a data frame.
# Use mutate() from the dplyr package:
```{r}
data4 <- dataset
head(data4 %>% mutate(n_student=n_student*2),n=20)
```
# Data normalization
# Loading all required libraries
```{r}
library(MASS)
library(dplyr)
library(gcookbook)
library(ggplot2)
```
# Box plot
```{r}
ggplot(dataset, aes(x=pretest, y=posttest))+geom_boxplot()
```
# Summarizing Data by Groups
```{r}
 # Load MASS library for the dataset data set
library(MASS)
library(dplyr)
library(tidyr)
```
```{r}
summarise(dataset, math.score = mean(pretest))
```

```{r}
ggplot(dataset,aes(x=pretest, fill=n_student, y=posttest))+geom_col(position ="dodge")
```

# Creating a Scatter Plot
```{r}
plot(dataset$pretest, dataset$posttest)
# Scatter plot with base graphics
```

# Scatter plot with ggplot 2
```{r}
library(ggplot2)
ggplot(dataset, aes(x = n_student, y = pretest)) +geom_point()

ggplot(data = NULL, aes(x =dataset$pretest, y=dataset$posttest )) +
geom_point()
```

# Creating a Line Graph
```{r}
plot(dataset$pretest,type = "l")
```

# Line graph with base graphics
```{r}
plot(head(dataset$pretest),head(dataset$posttest),  type = "l")
points(head(dataset$pretest),head(dataset$posttest))
```
# Line graph with ggplot() and With points added to ggplot()
```{r}
ggplot(head(dataset,n=20), aes(x = pretest, y = posttest)) +geom_line()
ggplot(head(dataset,n=20), aes(x = pretest, y = posttest)) +geom_line()+geom_point()
```
# Creating a Bar Graph
```{r}
barplot(head(dataset$n_student), names.arg = head(dataset$pretest))
```
# Generate a table of counts
```{r}
barplot(table(dataset$pretest))
```
# Laoding ggplot2() package
```{r}
library(ggplot2)
```
# Bar graph of values. This uses the dataset data frame, with the
# "pretest" column for x values and the "posttest" column for y values.
```{r}
ggplot(dataset, aes(x = pretest, y = posttest)) +geom_col()
```
# Convert the x variable to a factor, so that it is treated as discrete
```{r}
ggplot(dataset, aes(x = factor(pretest), y = posttest)) +geom_col()
```

# Bar graph of counts. This uses the dataset data frame, with the "pretest" column for
# x position. The y position is calculated by counting the number of rows for
# each value of pretest.
```{r}
ggplot(dataset, aes(x = pretest)) +
geom_bar()
# Bar graph of counts
ggplot(dataset, aes(x = factor(pretest))) +
geom_bar()
```
# Creating a Histogram
```{r}
hist(dataset$pretest)
# Specify approximate number of bins with breaks
hist(dataset$pretest, breaks = 10)
```
#ggplot2 histogram with default bin width (left); With wider bins (right)
```{r}
ggplot(dataset, aes(x = pretest)) +
geom_histogram()
ggplot(dataset, aes(x = pretest)) +
geom_histogram(binwidth = 4)
```
# Creating a Box Plot
```{r}
boxplot(head(dataset$n_student,n=500), head(dataset$pretest,n=500))
```
# Make box plots for multiple variables, by combining the variables with interaction()
```{r}
ggplot(head(dataset,n=100), aes(x = interaction(pretest, posttest), y = n_student)) +
geom_boxplot()
```
# Making a Basic Histogram
```{r}
ggplot(dataset, aes(x = pretest)) +geom_histogram()
```
```{r}
ggplot(dataset, aes(x= posttest)) +
geom_histogram(binwidth = 5, fill = "white", colour = "black")
```
# Map pretest to fill, make the bars NOT stacked, and make them semitransparent
```{r}
ggplot(dataset, aes(x = pretest, fill ='pretest',colour='red' )) +
geom_histogram(position = "identity", alpha = 0.4)
```
# Making a Density Curve
```{r}
ggplot(dataset, aes(x = n_student)) +geom_density()
# expand_limits() increases the y range to include the value 0
ggplot(dataset, aes(x = pretest)) +geom_line(stat = "density") +expand_limits(y = 0)
```
# Density curve with a smaller and larger value of adjust:
```{r}
ggplot(dataset, aes(x = posttest)) +geom_line(stat = "density") +
geom_line(stat = "density", adjust = .25, colour = "red") +
geom_line(stat = "density", adjust = 2, colour = "blue")
```
# This draws a blue polygon with geom_density(), then adds a line on top
# Density curve with wider x limits and a semitransparent fill (left); In two
# parts, with geom_density() and geom_line() (right)
```{r}
ggplot(dataset, aes(x = pretest)) +
geom_density(fill = "blue", alpha = .2) +
xlim(35, 105)
ggplot(dataset, aes(x = pretest)) +
geom_density(fill = "blue", alpha = .2, colour = NA) +
xlim(35, 105) +
geom_line(stat = "density")
```
# Density curve overlaid on a histogram
```{r}
ggplot(dataset, aes(x = pretest, y = ..density..)) +geom_histogram(fill = "cornsilk", colour = "grey60", size = .2) +geom_density() +xlim(35, 105)
```
# Making Multiple Density Curves from Grouped Data
```{r}
data5 <- dataset %>%
mutate(n_student = as.factor(n_student)) # Convert n_student to a factor
# Map n_student to colour
ggplot(data5, aes(x = pretest, colour = "red")) +geom_density()
# Map n_student to fill and make the fill semitransparent by setting alpha
ggplot(data5, aes(x = pretest, fill = pretest)) +geom_density(alpha = .3)
```
# Making a Frequency Polygon
```{r}
ggplot(dataset, aes(x=pretest)) +
geom_freqpoly()
ggplot(faithful, aes(x = waiting)) +      
geom_freqpoly(binwidth = 4)           #controlling bin width
```
# Divide the x-axis range into 15 bins
```{r}
binsize <- diff(range(dataset$pretest))/15
ggplot(dataset, aes(x = pretest)) +
geom_freqpoly(binwidth = binsize)
```
# Making a Basic Box Plot
```{r}
ggplot(dataset, aes(x = factor(n_student), y = pretest)) +
geom_boxplot()
```
# Box plot with narrower boxes (1)
# With smaller, hollow outlier points(2)
```{r}
ggplot(dataset, aes(x = factor(n_student), y = pretest)) +geom_boxplot(width = .5)
ggplot(dataset, aes(x = factor(n_student), y = pretest)) +geom_boxplot(outlier.size = 1.5, outlier.shape = 21)
```

```{r}
ggplot(dataset, aes(x = 1, y = pretest)) +geom_boxplot() +scale_x_continuous(breaks = NULL) +theme(axis.title.x = element_blank())
```

# Adding notches to a box plot to assess whether the medians are different.
```{r}
ggplot(dataset, aes(x = factor(n_student), y = pretest)) +geom_boxplot(notch = TRUE)
```
# Adding means to box plot
```{r}
ggplot(dataset, aes(x = factor(n_student), y = pretest)) +geom_boxplot() +stat_summary(fun = "mean", geom = "point", shape = 23, size = 3,
fill = "white")
```
# Making a Violin Plot
```{r}
data6 <- ggplot(dataset, aes(x = pretest, y = n_student))
data6+geom_violin()
```
# A violin plot with box plot overlaid on it
```{r}
data6+geom_violin() +geom_boxplot(width = .1, fill = "black", outlier.colour = NA) +
stat_summary(fun= median, geom = "point", fill = "white", shape = 21,
size = 2.5)
```
# A violin plot with tails
```{r}
data6+geom_violin(trim = FALSE)
```
# Violin plot with area proportional to number of observations
```{r}
data6 +geom_violin(scale = "count")
```
# Violin plot with
```{r}
# More smoothing
data6+geom_violin(adjust = 2)
# Less smoothing
data6 +geom_violin(adjust = .5)

```

# Making a Dot Plot
```{r}
# Save a modified data set that only includes of males data for marks that contains > 50
data7 <- dataset %>%
filter(gender == "Male" & pretest > 50)
```
```{r}
# Create a base ggplot object using `data7`, called `data7_p` (for data7 plot)
data8 <- ggplot(data7,aes(x=pretest))
data8+geom_dotplot()
```
# Dot plot with stackdir = “center” (1)
# With stackdir = “centerwhole”(2)
```{r}
data8 +
geom_dotplot(binwidth = .25, stackdir = "center") +
scale_y_continuous(breaks = NULL) +
theme(axis.title.y = element_blank())
data8 +
geom_dotplot(binwidth = .25, stackdir = "centerwhole") +
scale_y_continuous(breaks = NULL) +
theme(axis.title.y = element_blank())
```
# Making Multiple Dot Plots for Grouped Data
```{r}
ggplot(dataset, aes(x = gender, y = n_student)) +
geom_dotplot(binaxis = "y", binwidth = .5, stackdir = "center")
```
# Dot plot overlaid on box plot
```{r}
ggplot(dataset, aes(x = gender, y = n_student)) +
geom_boxplot(outlier.colour = NA, width = .4) +
geom_dotplot(binaxis = "y", binwidth = .5, stackdir = "center", fill = NA)

```
# Dot plot next to box plot
```{r}
ggplot(dataset, aes(x = pretest, y = n_student)) +
geom_boxplot(aes(x = as.numeric(pretest) + .2, group = pretest), width = .25) +geom_dotplot(aes(x = as.numeric(pretest) - .2, group = pretest),binaxis = "y",binwidth = .5,stackdir = "center") +scale_x_continuous(breaks = 1:nlevels(dataset$pretest),labels=levels(dataset$pretest)
)
```
# Making a Density Plot of Two-Dimensional Data
```{r}
# Save a base plot object
 ggplot(dataset, aes(x = pretest, y = posttest))+geom_point() +stat_density2d()
# Contour lines, with "height" mapped to color
ggplot(dataset, aes(x = pretest, y = posttest))+geom_point() +stat_density2d()+stat_density2d(aes(colour = ..level..))

```
# With ..density.. mapped to fill (1)
# With points, and ..density.. mapped to alpha(2)
```{r}
# Map density estimate to fill color
ggplot(dataset, aes(x = pretest, y = posttest))+geom_point() +stat_density2d()+stat_density2d(aes(fill = ..density..), geom = "raster", contour = FALSE)
# With points, and map density estimate to alpha
ggplot(dataset, aes(x = pretest, y = posttest))+geom_point() +stat_density2d()+geom_point() +stat_density2d(aes(alpha = ..density..), geom = "tile", contour = FALSE)
```
# Plotting a Function
# # The data frame is only used for setting the range
```{r}
# The normal distribution
ggplot(data.frame(x = c(-3, 3)), aes(x = x))  + stat_function(fun = dnorm)
```
```{r}
# The t-distribution with df=2
ggplot(data.frame(x = c(-3, 3)), aes(x = x)) + stat_function(fun = dt, args = list(df = 2))

```




















